---
title: "Data 607 Project 1"
author: "Armenoush Aslanian-Persico"
output:
  html_document:
    highlight: tango
    theme: cerulean
  pdf_document: default
---

#Load packages.
```{r}
library(knitr)
library(stringr)
library(tidyr)
library(dplyr)
```

#Import text file, identify delimiters, skip first 4 rows.
```{r}
chess<-read.table("tournamentinfo.txt", sep="|", fill=TRUE, stringsAsFactors = FALSE, skip=4)
#str(chess) #testing
#head(chess) #testing
```

#Combine observations across two rows.
```{r}
namesrow <- chess[seq(1,nrow(chess),by=3),]
#head(namesrow) #testing
staterow <- chess[seq(2,nrow(chess),by=3),]
#head(staterow) #testing
chess <- cbind(namesrow, staterow)
#head(chess) #testing
``` 

#Discard unneeded columns.
```{r}
chess <- subset(chess[, c(1:10,12,13)])
#head(chess) #testing
```

#Use regular expressions to extract player ID and pre-game rating.
```{r}
chess <- extract(chess, V2.1, c('PID', 'PreRating'), '(.[0-9]{8}) / R: (.[0-9]{3,4})')
#head(chess) #testing
```

#Use regular expressions to split each round into round number and opponent number.
```{r}
chess <- extract (chess, V4, c('Game1', 'Opp1'), '^([A-Z]{1})* +(.[0-9]{0,2})')
chess <- extract (chess, V5, c('Game2', 'Opp2'), '^([A-Z]{1})* +(.[0-9]{0,2})')
chess <- extract (chess, V6, c('Game3', 'Opp3'), '^([A-Z]{1})* +(.[0-9]{0,2})')
chess <- extract (chess, V7, c('Game4', 'Opp4'), '^([A-Z]{1})* +(.[0-9]{0,2})')
chess <- extract (chess, V8, c('Game5', 'Opp5'), '^([A-Z]{1})* +(.[0-9]{0,2})')
chess <- extract (chess, V9, c('Game6', 'Opp6'), '^([A-Z]{1})* +(.[0-9]{0,2})')
chess <- extract (chess, V10, c('Game7', 'Opp7'), '^([A-Z]{1})* +(.[0-9]{0,2})')
```

#View result and structure.
```{r}
knitr::kable(head(chess))
str(chess)
```

#Change required character columns to numeric, view structure.
```{r}
chess[, c(1,3,5,7,9,11,13,15,17,19,20)] <- sapply(chess[, c(1,3,5,7,9,11,13,15,17,19,20)], as.numeric)
str(chess)
```

#Create lookup table of all players and their pre-game rating, view structure.
```{r}
tblookup <- subset(chess[, c(1,20)])
str(tblookup)
```

#Perform lookup for each opponent by their opponent number.

```{r}
combo<- merge(chess, tblookup, by.x="Opp1",by.y="V1", sort=FALSE)
combo<- merge(combo, tblookup, by.x="Opp2",by.y="V1", sort=FALSE)
combo<- merge(combo, tblookup, by.x="Opp3",by.y="V1", sort=FALSE)
combo<- merge(combo, tblookup, by.x="Opp4",by.y="V1", sort=FALSE)
combo<- merge(combo, tblookup, by.x="Opp5",by.y="V1", sort=FALSE)
combo<- merge(combo, tblookup, by.x="Opp6",by.y="V1", sort=FALSE)
combo<- merge(combo, tblookup, by.x="Opp7",by.y="V1", sort=FALSE)
```

#Discard unneeded columns, rename columns, view result.
```{r}
combo <- subset(combo[, c(8:10,18:27)])
colnames(combo) <- c("Order","Player","Points","State","Player ID","PreRating","Opp1PR","Opp2PR","Opp3PR","Opp4PR","Opp5PR","Opp6PR","Opp7PR")
knitr::kable(combo)
```

#Add column for average opponent pre-game rating, view result.
```{r}
combo <- mutate(combo, OppMean = round(rowMeans(combo[,7:13]),0))
knitr::kable(combo)

```

#Discard unneeded columns, view result.
```{r}
results <- subset(combo[,c(1:6,14)])
knitr::kable(results)
```

#Export as CSV.
```{r}
write.table(results, "chessresults.txt")
```

#This can be done more easily.